---
title: "Natural_Language_Processing_Project"
output:
  pdf_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```

## Initialization

The initial step that loads the required libraries and downloads the data sets if
not all read on file. 
```{r initial}
library(tidyverse)
library(tidytext)
library(stopwords)


#downloads the corpus files, profanity filter and English dictionary

url <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
url2 <- "https://www.freewebheaders.com/download/files/facebook-bad-words-list_comma-separated-text-file_2021_01_18.zip"
url3 <- "https://raw.githubusercontent.com/dwyl/english-words/master/words_alpha.txt"
if(dir.exists("~/R/Capestone/data/") == FALSE){
       dir.create("~/R/Capestone/data/")}

if(file.exists("~/R/Capestone/data/data.zip") == FALSE|
   file.exists("~/R/Capestone/data/prof.zip")==FALSE|
   file.exists("~/R/Capestone/data/diction.txt")==FALSE){
        download.file(url,destfile = "~/R/Capestone/data/data.zip")
        download.file(url2,destfile = "~/R/Capestone/data/prof.zip")
        download.file(url3,destfile = "~/R/Capestone/data/diction.txt")
        setwd("~/R/Capestone/data/")
        unzip("~/R/Capestone/data/prof.zip")
        unzip("~/R/Capestone/data/data.zip")
        setwd("~/R/Capestone")
}
```

## Create corpus

At this stage the files are open and joined to create a corpus for the project. 
The Corpus is so large and requires some much ram that a sample of 20% is taken. 
```{r corpus}

blog <- read_lines("~/R/Capestone/data/final/en_US/en_US.blogs.txt")
news <- read_lines("~/R/Capestone/data/final/en_US/en_US.news.txt")
twitter <- read_lines("~/R/Capestone/data/final/en_US/en_US.twitter.txt")
blog <- tibble(text = blog) 
news <- tibble(text = news)
twitter <- tibble(text = twitter)

set.seed(90210)
corpus <- full_join(blog,twitter) %>% full_join(news) %>% 
        slice_sample(prop = 0.10) 
```

## Corpus filtering

Here the corpus filter is created to remove profanity and any word that is  not 
in the English dictionary. 
```{r filter}
prof <- read_lines("~/R/Capestone/data/facebook-bad-words-list_comma-separated-text-file_2021_01_18.txt")[15]
prof <- prof %>% str_split(", ") %>% flatten %>% unlist
prof <- tibble("word" = prof)

english <- read_lines("~/R/Capestone/data/diction.txt")
english <- tibble("word" = english[!english==""])
```

## Vocabulary 

A vocabulary of words is created from the unique words with the applied filters.
It also includes the removal of stop words which are very common words that provide
little value to predictive models. 
```{r vocab}
#clean up ram
rm(blog,news,twitter)
voc <- english %>% anti_join(prof) %>% anti_join(stop_words)

unigram <- corpus %>% unnest_tokens(ngram, text, token = "ngrams", n = 1) %>%
        semi_join(voc, by = c("ngram"="word")) 
#decreases the voc size
voc <- tibble(word = unigram$ngram)
```


## Out of Vocabulary 

To model out of vocabulary words we take a sampling of the least frequent unigrams
and change them to the character "<unk>". If a word is tested that isn't in the vocabulary
for the corpus, the quantity will be converted to "<unk>".
```{r OOV}
#OOV 1% of the least likely unigrams
unk <- unigram %>%
        count(ngram) %>%
        filter(n==1) %>%
        slice_sample(prop = 0.005) 
unigram[unigram$ngram %in% unk$ngram,]$ngram <- "<unk>"
unigram <- count(unigram, ngram)
```

There are still many word that only appear once (`r sum(unigram$n==1)`) so it
useful to remove these words as the add little to prediction value of the models
but significantly to the time and memory requirement. This is mainly due to the 
filtering. 

```{r vocabclean}

remove <- unigram$n==1
unigram <- unigram[!remove,]
voc <- tibble(word = unigram$ngram)
```
## Bigrams

The bigrams are created and then split into individual words which can be filtered
by the vocabulary list. 
```{r bigram}
bigram <- corpus %>% unnest_tokens(ngram, text, token = "ngrams", n = 2) %>%
        separate(ngram, c("word1", "word2"))
bigram$word1[bigram$word1 %in% unks$ngram] <- "<unk>"
bigram$word2[bigram$word2 %in% unks$ngram] <- "<unk>"
bigram <- bigram %>% filter(word1 %in% voc$word, word2 %in% voc$word) %>%
        count(word1, word2)
```

## Trigrams

Likewise, the trigrams are created in a similar manner
```{r trigrams}
trigram <- corpus %>% unnest_tokens(ngram, text, token = "ngrams", n = 3) %>%
        separate(ngram, c("word1", "word2", "word3"))
trigram$word1[trigram$word1 %in% unks$ngram] <- "<unk>"
trigram$word2[trigram$word2 %in% unks$ngram] <- "<unk>"
trigram$word3[trigram$word3 %in% unks$ngram] <- "<unk>"
trigram <- trigram %>% 
        filter(word1 %in% voc$word, word2 %in% voc$word, word3 %in% voc$word) %>%
        count(word1, word2, word3)
```

## Testing

A tibble is made for model testing, the test set itself is from one of the quizzes
```{r testdata}
dist <- tibble(word = voc$word)
input <- tibble(text = c("When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd",
                         "Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his",
                         "I'd give anything to see arctic monkeys this",
                         "Talking to your mom has the same effect as a hug and helps reduce your",
                         "When you were in Holland you were like 1 inch away from me but you hadn't time to take a",
                         "I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the",
                         "I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each",
                         "Every inch of you is perfect from the bottom to the",
                         "I’m thankful my childhood was filled with imagination and bruises from playing",
                         "I like how the same people are in almost all of Adam Sandler's")) %>%
        mutate(line = row_number())
input <- unnest_tokens(input, word, text)
input <- split(input, input$line) 

w <- vector(length = 3)
w[1] <- input$'1'[nrow(input$'1')-1,]$word
w[2] <- input$'1'[nrow(input$'1'),]$word
```

## Modeling
In this section, the different models are examined

### Maximum Likliness Estimate (MLE)
```{r MLE}
temp <- trigram %>% filter(word1 == w[1]) %>% filter(word2 == w[2])
MLEdf <- tibble(word=temp$word3, MLE=temp$n/sum(temp$n))
dist <- left_join(dist, MLEdf)
dist[is.na(dist$MLE),]$MLE <- 0
```

### Add one smoothing
```{r Addone}
ADDdf <- tibble(word=temp$word3, ADD=temp$n)
dist <- left_join(dist, ADDdf)
dist[is.na(dist$ADD),]$ADD <- 0
dist$ADD <- dist$ADD + 1
dist$ADD <- dist$ADD / sum(dist$ADD)
```

### Good Turing
```{r GT}
Nr <- count(temp, n) %>% add_row(n = 0, nn = 0)  %>% 
        arrange(n)
Nr %<>% mutate(c= 0) %>% mutate(Gt = 0)
total <- sum(Nr$nn*Nr$n)
Nr$Gt[Nr$n==0] <- Nr$nn[Nr$n==1] / total
Nr$nn[Nr$n==0] <- nrow(dist) - total
for (i in 2:nrow(Nr)) {
        Nr$c[i] <-  (Nr$n[i]+1)*Nr$nn[i+1]/Nr$nn[i]
        Nr$c[i][is.na(Nr$c[i])] <- 0
        Nr$Gt[i] <-  Nr$c[i] / total
}
Nr$Gt[Nr$n==max(Nr$n)] <- 1 - sum(Nr$Gt)
Nr$Gt <- Nr$Gt/ Nr$nn
Gt <- tibble(word=voc$word)
Gt <- left_join(Gt, select(temp, word3, n), by = c("word" = "word3"))
Gt$n[is.na(Gt$n)] <- 0 
Gt <- left_join(Gt, select(Nr, n, Gt))
dist <- left_join(dist, select(Gt, word, Gt))
```

### Absolute Discounting
```{r AD}
ADdf <- voc %>% left_join(select(temp, word3, n), by = c("word" = "word3")) 
ADdf$n[is.na(ADdf$n)] <- 0
d <- 0.5
temp2 <- trigram %>% filter(word1 == w[1]) %>% 
        count(word3) %>% 
        mutate(p3 = n / sum(n))

ADdf <- mutate(ADdf, p = n /sum(n), p1 = ((n-d) > 0) * (n-d) / sum(n),
               p2 = d/sum(n)* nrow(temp))
ADdf <- left_join(ADdf, select(temp2, word3, p3), by = c("word" = "word3" ))
ADdf$p3[is.na(ADdf$p3)] <- 0
ADdf <- mutate(ADdf, AD = p1+p2*p3)
```

### Kneser Ney

## Results

```{r results}
dist %>% summarise_at(c("MLE", "ADD", "Gt"), max)
```





